---
title: "Reproducible Research - Peer Assessment 1"
author: "Hoang Anh Tuan"
date: "Sunday, May 17, 2015"
---

##Loading and preprocessing the data

###Loading data
```{r}
setwd("D:\\Projects\\training\\DataScience Specilization Certificate\\Reproducible Research")
activity <- read.csv("activity.csv", header=TRUE,sep=",",na.strings="NA", colClasses=c("numeric", "character", "numeric"))
```

###Convert date time data to the corresponding data type
```{r}
activity$hour <- activity$interval %/% 100
activity$minute <- activity$interval %% 100
activity$datetime <- strptime(paste(activity$date, activity$hour, activity$minute), format="%Y-%m-%d %H %M")
activity$date <- as.Date(activity$date, format="%Y-%m-%d")
```

##What is mean total number of steps taken per day?
###Calculate the total number of steps taken per day
```{r, echo=TRUE, warning=FALSE}
data1 <- aggregate(steps ~ date, activity, sum)
```

###Make a histogram of the total number of steps taken each day
```{r, echo=TRUE, warning=FALSE}
library(ggplot2)
qplot(steps, data = data1)
```

###Calculate and report the mean and median of the total number of steps taken per day
```{r, echo=TRUE, warning=FALSE}
mean(data1$steps)
median(data1$steps)
```

##What is the average daily activity pattern?

###Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r, echo=TRUE, warning=FALSE}
data2 <- aggregate(steps ~ interval, activity, mean)
plot(data2$interval, data2$steps, type="l", xlab= "Interval", ylab= "Number of steps", col="green" , lwd=2)
```

###Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r, echo=TRUE, warning=FALSE}
with(data2, interval[steps == max(steps)])
```

##Imputing missing values

```{r, echo=TRUE, warning=FALSE}
sum(is.na(activity$steps))
newActivity <- activity
```

###Missing value at an interval is filled by the average value of steps at this interval

```{r, echo=TRUE, warning=FALSE}
newActivity$steps[which(is.na(newActivity$steps) & match(data2$interval, newActivity$interval))] <- data2$steps
```

```{r, echo=TRUE, warning=FALSE}
data3 <- aggregate(steps ~ date, newActivity, sum)
qplot(steps, data = data3)
mean(data3$steps)
median(data3$steps)
```

##Are there differences in activity patterns between weekdays and weekends?

```{r, echo=TRUE, warning=FALSE}
newActivity$WEND <- factor(newActivity$datetime$wday == 0 | newActivity$datetime$wday == 6,levels=c(TRUE,FALSE),labels=c("weekend","weekday"))
par(mfrow=c(2,1))
par(mar = c(2, 2, 1, 2))

weekendAct <- aggregate(steps ~ interval, data=subset(newActivity, WEND == "weekend"), mean)
weekdayAct <- aggregate(steps ~ interval, subset(newActivity, WEND == "weekday"), mean)
plot(weekendAct$interval, weekendAct$steps, type="l", main="Weekend", xlab= "Interval", ylab= "Number of steps", col="blue" , lwd=2)
plot(weekdayAct$interval, weekendAct$steps, type="l", main="Weekday", xlab= "Interval", ylab= "Number of steps", col="blue" , lwd=2)
```